---
title: "Making adjustments"
author: "Anders H. Jarmund"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.asp = 0.7,
  fig.width = 12,
  fig.align = "center",
  cache = FALSE,
  external = FALSE
)
```

# Introduction

<script src="js/hideOutput.js"></script>
  <style>
  .showopt {
    background-color: #004c93;
      color: #FFFFFF;
      width: 150px;
    height: 20px;
    text-align: center;
    vertical-align: middle !important;
    float: left;
    font-family: sans-serif;
    border-radius: 8px;
  }

.showopt:hover {
  background-color: #dfe4f2;
    color: #004c93;
}

.showopttext::before{
  content: "Show/Hide Source"
}

pre.plot {
  background-color: white !important;
}
</style>

## Load data

```{r setup}
# library("inpreg")
devtools::load_all("C:/Users/anderhja/Lokal_folder/inpreg2")
obj <- inpreg()
```

## Find duplicates

Based on study + id + time

```{r}
obj$set_batch(batch = NULL)
df_dup <- obj$find_duplicates()
obj$plot_duplicates(ref = "2", scales = "free")
```

## Adjusting batch effects

```{r}
obj$set_batch(batch = c("1", "2", "4"))
obj$set_type(type = "X")
g_pre <- obj$plot_pca()
obj$adjust_batch_effect(types = "X", ref = "2")
g_post <- obj$plot_pca()
ggpubr::ggarrange(g_pre, g_post, labels = "AUTO")


df <- obj$df_cmp(types = "X", wide = FALSE)[, .(value = median(value, na.rm = TRUE)), by = c("bp", "variable", "adjusted")]
ggplot(df, aes(bp, value, color = adjusted)) + geom_point() + facet_wrap(~variable, scales = "free_y") + coord_flip() + obj$my_theme

df <- obj$df_cmp(types = "X", wide = FALSE)[, .(value = median(value, na.rm = TRUE)), by = c("bp", "variable", "study", "adjusted")]
ggplot(df, aes("study", value, color = adjusted)) + geom_point() + facet_wrap(~variable, scales = "free_y") + coord_flip() + obj$my_theme
```

## Adjusting plate effects

```{r}
obj$set_batch(batch = c("1","2"))
obj$set_type(type = "X")
g_pre <- obj$plot_pca(color = "plate")

obj$set_batch(batch = c("2"))
obj$adjust_plate_effect(types = "X", ref_study = "overtid", ref_time = 1)

obj$set_batch(batch = c("1"))
obj$adjust_plate_effect(types = "X", ref_study = "pregmet1", ref_time = 1)
obj$adjust_plate_effect(types = "X", ref_study = "normalflow", ref_time = 1)

obj$set_batch(batch = c("1","2"))
g_post <- obj$plot_pca(color = "plate")
ggpubr::ggarrange(g_pre, g_post, labels = "AUTO")


df <- obj$df_cmp(types = "X", wide = FALSE)
ggplot(df, aes(plate, value, color = adjusted)) + geom_point() + facet_wrap(~variable, scales = "free_y")

df <- obj$c_o_v(df = obj$df_cmp(types = "X", wide = FALSE), by = c("batch", "study", "adjusted", "time"))
ggplot(df, aes(paste(study, time), cov, color = adjusted)) + geom_point() + facet_wrap(~variable, scales = "free_y") + coord_flip() + obj$my_theme
```

## Getting cytokine concentrations

```{r}
# obj$set_batch(batch = c("2"))
# obj$get_concentrations(ref = "2")
# obj$plot_standards()
```
