---
title: "Diagnostics"
author: "Anders H. Jarmund"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.asp = 0.7,
  fig.width = 12,
  fig.align = "center",
  cache = FALSE,
  external = FALSE
)
```

# Introduction

<script src="js/hideOutput.js"></script>
  <style>
  .showopt {
    background-color: #004c93;
      color: #FFFFFF;
      width: 150px;
    height: 20px;
    text-align: center;
    vertical-align: middle !important;
    float: left;
    font-family: sans-serif;
    border-radius: 8px;
  }

.showopt:hover {
  background-color: #dfe4f2;
    color: #004c93;
}

.showopttext::before{
  content: "Show/Hide Source"
}

pre.plot {
  background-color: white !important;
}
</style>


## Load data

```{r setup}
# library("inpreg")
devtools::load_all("C:/Users/anderhja/Lokal_folder/Processing-cytokine-data-in-R/")
obj <- inpreg()
```

## PCA

We are typically using PCA to evaluate if batch or plate is the major cause of differences between the samples. PCA can done with

```{r}
# Plot patient samples only
obj$set_type(type = "X")

# Color by batch
obj$plot_pca(color = "batch")
```

You can highlight specific batches by using (here: batch 1, plate 1 and 2):

```{r}
obj$plot_pca(highlight = c("b1-p1", "b1-p2"))
```

## Interactive plots

To get interactive plots with `ggplotly`:

```{r}
# It remembers this form previously but repeating just to be sure
obj$set_type(type = "X")
plotly::ggplotly(
  obj$plot_pca(highlight = c("b1-p1", "b1-p2"))
)
```

## Exclude variables

If you want to exclude variables from the analysis, use

```{r}
obj$rm_variables(c("IL_10", "RANTES", "IL_5"))
obj$plot_boxplot()
obj$plot_pca(color = "batch")
obj$rm_variables(NULL) # Include all variables again
```

## Plot control samples

By default, all types are shown.

```{r}
# Show only batch 2
obj$set_batch(batch = "2")

# Plot quality control samples
obj$plot_qc()
```

You can reduce the number by specifying what kind of samples you want: `c("blank", "internkontroll", "serumkontroll", "internstandard")`

```{r}
obj$plot_qc(types = c("blank", "serumkontroll"), scales = "free_x")
```

## Plot medians

Plot the medians and 25. and 75. percentiles

```{r}
obj$set_type("X")
obj$plot_time()
```

## Plot time development

Plot samples by participants, over time

```{r}
obj$plot_medians()
```

## Plot standard curves

```{r}
obj$set_batch(batch = "2")
obj$plot_standards(highlight = "b2-p9")
```

## Find duplicates

Based on study + id + time

```{r}
obj$set_batch(batch = NULL)
df_dup <- obj$find_duplicates()
obj$plot_duplicates(ref = "2", scales = "free")
```
