---
title: "Get Started"
author: "Anders H. Jarmund"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%", 
  fig.asp = 0.7,
  fig.width = 12,
  fig.align = "center",
  cache = FALSE,
  external = FALSE
)
```

# Introduction

<script src="js/hideOutput.js"></script>
<style>
.showopt {
  background-color: #004c93;
  color: #FFFFFF; 
  width: 150px;
  height: 20px;
  text-align: center;
  vertical-align: middle !important;
  float: left;
  font-family: sans-serif;
  border-radius: 8px;
}

.showopt:hover {
    background-color: #dfe4f2;
    color: #004c93;
}

.showopttext::before{
  content: "Show/Hide Source"
}

pre.plot {
  background-color: white !important;
}
</style>

## Loading data

The idea with this package is to simplify the normalization and pre-processing of multi-batch cytokine data.

```{r setup}
# library("inpreg")
# devtools::install_version("nCal", version = "2021.9-12", repos = "http://cran.us.r-project.org")
devtools::load_all("C:/Users/anderhja/Lokal_folder/inpreg2")
obj <- inpreg(
  filenames = list(
                       FI = "C:/Users/anderhja/Lokal_folder/git/fluorescens_alle.csv",
                       biorad = "C:/Users/anderhja/Lokal_folder/git/raw_alle.csv",
                       expected = "C:/Users/anderhja/Lokal_folder/git/fluorescens_expected.csv"
                     )
)
```

Note: By default, all values are log-scaled. You can now make a simple plot of all FI values

```{r}
obj$plot_boxplot()
```

If you prefer biorad concentrations rather than FI values, you haveto specify this when you create the object:

```{r}
# library("inpreg")
obj_biorad <- inpreg(use_FI = FALSE)
obj$plot_boxplot()
```

## Selecting specific sample types and batches

Note that _all_ samples are used by default. You can select specific sample types by selecting one of `types = c("X", "S", "C")`; `X` is serum, `S` is the standards and `C` refers to control samples. You can also redefine the `x_axis` of the boxplot. For example:

```{r}
# Use serum samples only
obj$set_type(type = "X")

# USe `study` as x axis
obj$plot_boxplot(x_axis = "study")
```

By default, all batches are included (note: those with negative batch (or plate!) number are excluded). To select specific batches, list them as follows:

```{r}
obj$set_batch(batch = c("1", "2"))
obj$plot_boxplot()
```

And to select all again:

```{r}
obj$set_batch(batch = NULL)
```

## Count missing values

Note: This is based on biorad data, not FI (as we are not missing any FI data). Since we will estimate our own values, are the biorad data only indicative

By batch and plate (shorthand: `bp`):

```{r}
obj$plot_missing(x_axis = "bp")
```

By study in batch 2:

```{r}
obj$set_batch(batch = "2")
obj$plot_missing(x_axis = "study")
```

## Interactive plots

To get interactive plots with `ggplotly`:

```{r}
plotly::ggplotly(
  obj$plot_missing(x_axis = "study")
)
```

## Plot medians

Plot the medians and 25. and 75. percentiles

```{r}
obj$set_type("X")
obj$set_batch(batch = "2")
plotly::ggplotly(
  obj$plot_medians(x_axis = "study")
)
```
